---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library("tidyverse")
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
spotify_song <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify_song |>
summarise(
    mean_danceability = mean(danceability),
    sd_danceability   = sd(danceability)
  )
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
spotify_song |>
  ggplot(aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "white") +
  labs(title = "Histogram of Danceability", x = "Danceability", y = "Count")

```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
spotify_song |>
  ggplot(aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "white") +
  geom_vline(aes(xintercept = mean(danceability)),
             color = "red", linetype = "dashed", size = 1) +
  labs(title = "Histogram of Danceability", x = "Danceability", y = "Count")

```


### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
m <- mean(spotify_song$danceability, na.rm = TRUE)
s <- sd(spotify_song$danceability, na.rm = TRUE)

spotify_song |>
  ggplot(aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "white") +
  geom_vline(xintercept = m, color = "red", size = 1.2) +          # Mean
  geom_vline(xintercept = m - s, color = "darkgreen", linetype = "dashed", size = 1) + # Mean - 1 SD
  geom_vline(xintercept = m + s, color = "darkgreen", linetype = "dashed", size = 1) + # Mean + 1 SD
  labs(title = "Histogram of Danceability with Mean ±1 SD",
       x = "Danceability", y = "Count")
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.
The danceability scores in Spotify’s dataset are slightly skewed toward higher values, meaning most songs are fairly easy to dance to. The mean shows the average groove, and the standard deviation tells us that most songs cluster around this average.

Story pitch:
"Most songs on Spotify aren’t just for listening — they’re made to move us. With most tracks near the average danceability, the dataset reveals the pulse of modern music we want to feel the beat and songs that have the most dance vibe to it."


### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
