---
title: "HW7_MandMs"
date: 10/1/2025
---


In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library("tidyverse")
```


### Task 2: Load an M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
MMs <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
# Set the random seed for reproducibility
set.seed(1234)

# Sample 5 rows from the dataset
sample5 <- MMs |>
  sample_n(5)

# Sample 10 rows from the dataset
sample10 <- MMs |>
  sample_n(10)


```


### Task 4: Calculate the mean, median and standard deviation of red M&Ms from each of the samples.
```{r}

red_stats <- MMs |>
  group_by(total) |>
  summarize(
    mean_red = mean(red),
    median_red = median(red),
    sd_red = sd(red,)
  )

red_stats

```


### Task 5: Calculate the mean, median and standard deviation of red M&Ms from the population

```{r}
red_mandm <- MMs|>
summarize(
    mean_red = mean(red),
    median_red = median(red),
    sd_red = sd(red,)
  )

```

### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Include a headline. Use a consistent binwidth.

```{r}
MMs |>

# Create histograms for sample5, sample10, and the population
ggplot() +
  geom_histogram(data = sample5, aes(x = red), binwidth = 1, fill = "red", alpha = 0.6) +
  geom_histogram(data = sample10, aes(x = red), binwidth = 1, fill = "blue", alpha = 0.6) +
  geom_histogram(data = MMs, aes(x = red), binwidth = 1, fill = "green", alpha = 0.6) +
  labs(
    title = "Distribution of Red M&Ms: Sample 5, Sample 10, and Population",
    x = "Number of Red M&Ms",
    y = "Frequency"
  ) +
  theme_minimal()

```


### Task 7: Answer questions


Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?
The population statistics are most accurate, while larger samples (like sample10) better reflect the population than smaller ones.

What does this exercise tell you about the relationship between sample size and estimate accuracy?
Larger samples give more accurate estimates of the population values.

How does random sampling introduce variation into statistical estimates?
Random sampling introduces variation because each sample contains different data points.

If you repeated this process, would the samples always be identical? Why or why not?
No, samples wouldn’t be identical since random selection changes which rows are chosen each time.


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
