---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library2("tidyverse")

```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
MandM_data <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")

```


### Task 3: Sample 5 rows of data from the dataset
```{r}
sampled_rows <- MandM_data |>
  sample_n(5)

sampled_rows

```


### Task 4: Calculate the mean age and standard deviation of the sample, using either the blue or white number of M&Ms
```{r}
# Filter for rows where red > 0
red_sample <- MandM_data |> 
  filter(red > 0)

# Calculate mean and sd of the age column
mean_age_red <- mean(red_sample$age, na.rm = TRUE)
sd_age_red   <- sd(red_sample$age, na.rm = TRUE)

mean_age_red
sd_age_red




```


### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4
```{r}
# Calculate mean and standard deviation of age for the full population
mean_age_population <- mean(MandM_data$age, na.rm = TRUE)
sd_age_population   <- sd(MandM_data$age, na.rm = TRUE)

mean_age_population
sd_age_population

```
### Task 6: Compare the sample and population statistics
Compare the two sets of means and standard deviations. Are they similar?

Yes it gave me NA_real both times
### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

The result came out as N/A, likely because the dataset contained missing or incomplete values. This shows that random sampling and analysis depend on having valid data — if parts of the dataset are missing, it can prevent accurate calculations or meaningful comparisons.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
