---
title: "HW9_TestScores"
Name: Derek Willis/Rob Wells
Date: 10/21/2025
---

name: Sam Ford


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Did a New Reading Program Lead to Better Scores?

The superintendent recently claimed that a new reading program has improved third-grade reading scores across the school district.

Before the program, third-grade students in the district averaged 72.6 points on standardized reading tests with a standard deviation of 4.8 points.

After implementing the program for one semester, you collected scores from 12 randomly selected classrooms:
74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75

As a journalist, you need to determine: **Is there statistical evidence that reading scores have actually improved?**

## Task 1: Organize your data and initial assessment

Before you can run this codeblock, you will need to fill in a value where it says REPLACE_ME. That value can be found in the introduction.

```{r}
# Known information about reading scores before the new program
prior_mean <- 72.6  # average score
prior_sd <- 4.8     # standard deviation

# Reading scores after implementing the new program (12 classrooms)
new_scores <- c(74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75) # Replace with the actual scores

# Create a journalist-friendly dataset
score_data <- tibble(
  classroom = paste("Classroom", 1:12),
  reading_score = new_scores
)

# View the data
score_data
```

### Reflection Question 1:  (2 points)
Based on just looking at the score_data dataframe, have test scores improved? How can you tell? 
thet=y are higher because the average was 72.6 but all the new scores are higher than that.


## Task 2: Calculate key statistics  (2 points)

Like Task 1, you will need to replace values where it says REPLACE_ME before running any code.


```{r}
# Calculate statistics based on the new reading scores
new_stats <- score_data |> 
  summarise(
    mean = mean(72.6),
    sd = sd(4.8),
    n = n()
  )

new_stats
```

### Reflection Question 2: (3 points)
Looking at the mean and standard deviation of the new scores compared to the previous statistics, what initial observations can you make? Provide examples from the data in your answer. What questions might these statistics raise for your reporting?

The new average score is higher than before, suggesting students improved overall. With a standard deviation of 4.8, most students scored fairly close to the average, meaning performance was pretty consistent across the group. What specific changes led to this progress?Are students continuing to improve over time, or was this a one-time jump?

## Task 3: Create a column chart  (1 points)

As before, replace any values marked REPLACE_ME based the instructions.


```{r}
# STUDENT TASK: Choose an appropriate fill color for the bars
my_fill_color <- "royalblue" # Replace with a color name like "royalblue", "darkgreen", etc.

# Create a visualization comparing new scores to the previous average
score_data |> 
ggplot(aes(x = classroom, y = reading_score)) +
  geom_col(fill = my_fill_color, alpha = 0.8) +
  geom_hline(yintercept = prior_mean, color = "darkred", size = 1, linetype = "dashed") +
  annotate("text", x = 2, y = prior_mean - 1, 
           label = "Previous Average (72.6)",  hjust = -0.5, vjust= 2, fontface = "bold", color = "darkred") +
  labs(
    title = "Reading Scores After New Program Implementation",
    subtitle = "Horizontal line shows previous district average of 72.6 points",
    x = NULL,
    y = "Reading Test Score",
    caption = "Source: District Assessment Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Reflection Question 3: (3 points)
Examine the chart you created, and suggest a better title based on the results of the data, not a description.

Reading Scores with new program in each class

## Task 4: Perform a hypothesis test (2 points)

This is where we formally test the superintendent's claim that reading scores have improved. Fill in the REPLACE_ME values as needed, beginning with your hypotheses.

**Hypotheses:**
Null: The new reading program will hopefully improve numbers
Alternative: The new reading program will keep numbers the same

```{r}
# Set the significance level for your test
alpha_level <- 0.05 # Replace with the appropriate value

# Perform a one-sample t-test
# Since we want to know if scores improved (increased), we use a one-sided test (alternative = "greater")
t_test_result <- t.test(
  score_data$reading_score,
  mu = prior_mean,
  alternative = "greater"
)

# Display the results
t_test_result
```

### Reflection Question 4:  (3 points)
What does the p-value tell you, and what doesn't it tell you? How would you explain these results to a non-technical audience while maintaining accuracy?
The data suggests students really did improve, and it’s unlikely this happened by luck, though we can’t say for sure what caused it.


## Task 5: Interpreting the results for your news story (2 points)

Let's gather all of the important stats we'll need in one place, so we can look at the prior average, the new scores and the results of the t.test, including the confidence interval. Replace any values where it says REPLACE_ME.


```{r}
# Get the p-value
p_value <- t_test_result$p.value

# Calculate the 95% confidence interval
ci <- t.test(score_data$reading_score)$conf.int

# Create a tibble to display the key statistics for your story
story_stats <- tibble(
  `Previous average` = prior_mean,
  `New average` = mean(75.75),
  `Improvement` = mean(new_scores) - prior_mean,
  `Percent change` = round(((mean(new_scores) - prior_mean) / prior_mean) * 100, 1),
  `p-value` = p_value,
  `Lower bound` = ci[1],
  `Upper bound` = ci[2],
  `Confidence level` = "95%"
)

# Display the key statistics
story_stats
```

## Conclusion

### Reflection Question 5: (4 points)
Based on these statistics, what would be your headline and lead paragraph for this story? Is there evidence to support the superintendent's claim?

Students’ Reading Scores Show Real Improvement, Data Suggests”

New test results indicate that students’ average reading scores have risen significantly compared to last year. The increase is unlikely to be due to chance, suggesting real progress. While the data supports the superintendent’s claim of improvement, more evidence would be needed to confirm what caused the gains

### Reflection Question 6:  (1 point)
What metrics or outcomes beyond test scores might be important to track for assessing reading performance?

Other things that matter include how much students actually enjoy reading, how often they read on their own, and whether they can understand and talk about what they read. Teachers might also look at reading growth over time, class participation, and confidence with reading new material, not just the test numbers.
