---
title: "final_project"
output: html_document
date: "2025-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The data for this project comes from the Billboard Hot 100 Dataset, which tracks weekly chart performance for popular music in the United States. It includes metadata about songs such as title, artist, release year, chart ranking, peak position, and weeks on the chart.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
install.packages("tidyverse")
library(tidyverse)

billboard <- read_csv("Billboard Hot 100 Number Ones Database - Data.csv")
glimpse(billboard)
summary(billboard)


```

##3What the Dataset Contains

The Billboard Hot 100 data includes:

- **Song,Artist** (title, artist, producers)
- **Chart performance** (peak position, weeks on chart, year)
- **Date variables** (entry date, chart week)
- **Identifiers** (key notes, danciness, vocals)

To view summary statistics:

```{r}
summary(billboard)
```

To check the total rows/columns:

```{r}
dim(billboard)
```

# 4. Important Columns for This Analysis

For a project about “what makes a hit song,” the following variables matter most:

- **title** — identifies the song  
- **artist** — identifies the creator  
- **peak_position** — best chart rank (indicator of success)  
- **weeks_on_chart** — endurance of the song  
- **year** — needed for year-to-year trend analysis  
- **genre** — if included, helps analyze which genres perform well  

These variables allow us to measure success and compare across time periods.

---

# 5. What’s Missing From the Dataset

Some useful information that is not included or may be missing:

- **Marketing / promotion data**  
- **Streaming vs. radio vs. sales breakdown**  
- **Social media virality metrics**  

These limitations may restrict certain types of story ideas.

---

# 6. Basic Descriptive Statistics  
Using *weeks_on_chart* as the selected variable:

```{r}
billboard|>
 summarize(
    mean_weeks   = mean(`Weeks at Number One`, na.rm = TRUE),
    median_weeks = median(`Weeks at Number One`, na.rm = TRUE),
    min_weeks    = min(`Weeks at Number One`, na.rm = TRUE),
    max_weeks    = max(`Weeks at Number One`, na.rm = TRUE)
  )
```

These values show how long the average song lasts on the Hot 100 compared with the longest-lasting and shortest-lasting records.

---

# 7. Hypotheses and Newsworthy Questions

Below are hypotheses that can be tested using Billboard data:

### **Hypothesis 1:**  
Pop songs that reach the Top 10 have stronger chart performance than non-pop songs that reach the Top 10.

Null Hypothesis (H₀):
Among Top 10 hits, pop songs and non-pop songs have no difference in their chart performance (measured by total weeks on the Hot 100 or peak position).

Alternate Hypothesis (H₁):
Among Top 10 hits, pop songs stay on the Hot 100 longer or achieve better peak positions than non-pop Top 10 songs. 
### **Hypothesis 2:**  
More danceable songs perform better on the Billboard Hot 100.

Null Hypothesis (H₀):
Danceability has no relationship with chart performance (peak position or total weeks on the chart).

Alternate Hypothesis (H₁):
Songs with higher danceability scores achieve better chart performance — such as higher peak positions or more weeks on the Hot 100.

### **Hypothesis 3:**  
Hip-hop and R&B dominate the Hot 100 more in the streaming era.

Operational Definition of “dominate”:
Hip-hop and R&B songs make up more than 50% of all Hot 100 entries in a given period.

Time periods:

Streaming era: 2015–present

Pre-streaming era: 1980–2014 (or earlier based on your dataset)

Null Hypothesis (H₀):
Hip-hop/R&B songs do not make up a greater share of Hot 100 entries in the streaming era than before 2015.

Alternate Hypothesis (H₁):
Hip-hop/R&B songs make up a greater than 50% share of Hot 100 entries in the streaming era, showing increased dominance.

### **Hypothesis 4:**  
Songs that contain samples from older recordings perform better on the Billboard Hot 100 than songs that do not use samples.

Null Hypothesis (H₀):
There is no difference in chart performance between sampled songs and non-sampled songs. Sampling does not improve or worsen chart success.

Alternate Hypothesis (H₁):
Sampled songs achieve stronger chart performance — such as higher peak positions or more total weeks on the chart — compared to songs that do not include samples.
### **Hypothesis 5:**  
The number of weeks a song spends in the Top 10 has increased in the streaming era.

Null Hypothesis (H₀):
There is no difference in Top 10 duration before and after 2015.

Alternate Hypothesis (H₁):
Songs in the streaming era spend more weeks in the Top 10.

---
z score
```{r}
# Z-scores for Weeks at Number One
billboardz <- billboard |>
  mutate(
    z_weeks = (`Weeks at Number One` - mean(`Weeks at Number One`, na.rm = TRUE)) /
              sd(`Weeks at Number One`, na.rm = TRUE)
  )

# Explanation:
# "A positive z-score means a song stayed at #1 longer than usual,
# while a negative z-score means it didn’t last as long. 
# For example, a song with a z-score of 2 stayed on top much longer than most other hits. 
# This makes it easy to spot songs that were real chart dominators versus those that barely hung on at #1."

## T-test for Danceability

# Step 1: Remove NAs for Danceability and Weeks at Number One
billboard_clean <- billboard |>
  filter(!is.na(Danceability), !is.na(`Weeks at Number One`))

# Step 2: Create High vs Low Danceability groups based on median
billboard_clean <- billboard_clean |>
  mutate(
    dance_group = ifelse(Danceability >= median(Danceability, na.rm = TRUE),
                         "High Danceability",
                         "Low Danceability")
  )

# Step 3: Pull groups for testing
high_group <- billboard_clean |>
  filter(dance_group == "High Danceability") |>
  pull(`Weeks at Number One`)

low_group <- billboard_clean |>
  filter(dance_group == "Low Danceability") |>
  pull(`Weeks at Number One`)

# Step 4: Run the t-test
dance_t_test <- t.test(high_group, low_group, alternative = "greater")
dance_t_test



```
"The T-test compared songs with high danceability to those with low danceability in terms of how long they stayed at Number One on the Hot 100. A positive and significant result indicates that songs with higher danceability tend to remain at the top of the chart longer than less danceable songs. This suggests that danceability may play an important role in a song’s commercial success, potentially because more danceable songs are more likely to be streamed, shared, and played in clubs or social settings."

##data chart
```{r}
billboard |>
  group_by(run_type) |>
  summarize(
    avg_weeks = mean(`Weeks at Number One`, na.rm = TRUE)
  ) |>
  ggplot(aes(x = run_type, y = avg_weeks, fill = run_type)) +
  geom_col() +
  labs(
    title = "Average Weeks at #1: Long Run vs Short Run Songs",
    x = "Run Type",
    y = "Average Weeks at Number One"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
"The bar chart shows the average weeks at #1 for Long Run versus Short Run songs. Long Run hits clearly stay on top much longer, while Short Run hits drop off more quickly. It’s an easy way to see how not all #1 songs dominate the charts equally."

Story Idea:
Our analysis explores how danceability contributes to a song’s chart success on the Billboard Hot 100. Using T-tests to compare songs with high vs. low danceability, we found that more danceable songs tend to stay at Number One longer than less danceable songs. This suggests that danceability is a measurable factor influencing a song’s popularity and staying power, potentially amplified by streaming platforms, social media sharing, and club or party play.

Main Findings:
High-danceability songs remain at Number One significantly longer than low-danceability songs.Danceability may act as a predictor for commercial success, independent of genre or artist fame.Visualizations (bar charts and z-scores) make it clear which songs dominated the charts versus those that had brief peaks.

Potential Problems / Limitations:
Danceability data may not cover all songs or could include missing/incorrect values.Weeks at Number One does not capture songs that performed well overall but never reached #1.Other factors like marketing, social media virality, collaborations, or label support could also influence success and are not included in this analysis.

Reporting Plan:
Highlight the top high-danceability songs and their chart performance. Include bar charts comparing high vs. low danceability, and z-scores to show extreme performers.Discuss the broader cultural context: how danceable songs dominate TikTok, streaming playlists, and clubs. Interview music producers or DJs (optional) to provide expert commentary on why danceability matters. Conclude with actionable insights for music fans, industry analysts, or emerging artists.
