---
title: "final_project"
output: html_document
date: "2025-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The data for this project comes from the Billboard Hot 100 Dataset, which tracks weekly chart performance for popular music in the United States. It includes metadata about songs such as title, artist, release year, chart ranking, peak position, and weeks on the chart.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
install.packages("tidyverse")
library(tidyverse)

billboard <- read_csv("Billboard Hot 100 Number Ones Database - Data.csv")
glimpse(billboard)
summary(billboard)


```

##3What the Dataset Contains

The Billboard Hot 100 data includes:

- **Song,Artist** (title, artist, producers)
- **Chart performance** (peak position, weeks on chart, year)
- **Date variables** (entry date, chart week)
- **Identifiers** (key notes, danciness, vocals)

To view summary statistics:

```{r}
summary(billboard)
```

To check the total rows/columns:

```{r}
dim(billboard)
```

# 4. Important Columns for This Analysis

For a project about “what makes a hit song,” the following variables matter most:

- **title** — identifies the song  
- **artist** — identifies the creator  
- **peak_position** — best chart rank (indicator of success)  
- **weeks_on_chart** — endurance of the song  
- **year** — needed for year-to-year trend analysis  
- **genre** — if included, helps analyze which genres perform well  

These variables allow us to measure success and compare across time periods.

---

# 5. What’s Missing From the Dataset

Some useful information that is not included or may be missing:

- **Marketing / promotion data**  
- **Streaming vs. radio vs. sales breakdown**  
- **Social media virality metrics**  

These limitations may restrict certain types of story ideas.

---

# 6. Basic Descriptive Statistics  
Using *weeks_on_chart* as the selected variable:

```{r}
billboard|>
 summarize(
    mean_weeks   = mean(`Weeks at Number One`, na.rm = TRUE),
    median_weeks = median(`Weeks at Number One`, na.rm = TRUE),
    min_weeks    = min(`Weeks at Number One`, na.rm = TRUE),
    max_weeks    = max(`Weeks at Number One`, na.rm = TRUE)
  )
```

These values show how long the average song lasts on the Hot 100 compared with the longest-lasting and shortest-lasting records.

---

# 7. Hypotheses and Newsworthy Questions

Below are hypotheses that can be tested using Billboard data:

### **Hypothesis 1:**  
Songs that peak at Number 1 stay on the chart significantly longer than songs that peak below the Top 10.

### **Hypothesis 2:**  
Hit songs have become shorter over time (compare average song length by year, if length is available).

### **Hypothesis 3:**  
Hip-hop and R&B dominate the Hot 100 more in the streaming era (2015–present) compared to earlier decades.

### **Hypothesis 4:**  
Certain artists or record labels have disproportionate chart success (possible industry concentration story).

### **Hypothesis 5:**  
The number of weeks a song stays in the Top 10 has increased over time due to streaming.

---
z score
```{r}
billboardz <- billboard |>
  mutate(
    z_weeks = (`Weeks at Number One` - mean(`Weeks at Number One`, na.rm = TRUE)) /
              sd(`Weeks at Number One`, na.rm = TRUE)
  )
```
"A positive z-score means a song stayed at #1 longer than usual, while a negative z-score means it didn’t last as long. For example, a song with a z-score of 2 stayed on top much longer than most other hits. This makes it easy to spot songs that were real chart dominators versus those that barely hung on at #1."
##Ttest
```{r}

# Convert the column to numeric (fixes all errors)
billboard$`Weeks at Number One` <- as.numeric(billboard$`Weeks at Number One`)

# 1. Calculate the median
median_weeks <- median(billboard$`Weeks at Number One`, na.rm = TRUE)

# 2. Create Long Run vs Short Run groups
billboard <- billboard |>
  mutate(
    run_type = ifelse(`Weeks at Number One` >= median_weeks,
                      "Long Run",
                      "Short Run")
  )

# 3. Pull the groups
group1 <- billboard |>
  filter(run_type == "Long Run") |>
  pull(`Weeks at Number One`)

group2 <- billboard |>
  filter(run_type == "Short Run") |>
  pull(`Weeks at Number One`)

# 4. Run the t-test
t_test_result <- t.test(group1, group2)
t_test_result


```
"For the Ttest compared songs that stayed at #1 for a long time (“Long Run”) to those that didn’t stay as long (“Short Run”). The t-test shows a clear difference between the two groups: songs in the Long Run group stayed on the chart significantly longer than the Short Run songs. In other words, some #1 hits dominate the charts for weeks, while others hit the top but fall off more quickly. This confirms that there’s real variation in how long songs stay at the top spot"

##data chart
```{r}
billboard |>
  group_by(run_type) |>
  summarize(
    avg_weeks = mean(`Weeks at Number One`, na.rm = TRUE)
  ) |>
  ggplot(aes(x = run_type, y = avg_weeks, fill = run_type)) +
  geom_col() +
  labs(
    title = "Average Weeks at #1: Long Run vs Short Run Songs",
    x = "Run Type",
    y = "Average Weeks at Number One"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
"The bar chart shows the average weeks at #1 for Long Run versus Short Run songs. Long Run hits clearly stay on top much longer, while Short Run hits drop off more quickly. It’s an easy way to see how not all #1 songs dominate the charts equally."
